\thetitle

### 2021-05-25 writing/landscape.tex
fix small words, vspace for table on pg. 4
clarifying blame and boundaries + a small edit about narrow vs broad blame information

\end{description}
The mistaken comment in {\tt pack-lib} causes an impedance mismatch,
with which each of the three semantics deals differently.

Under {\it the Natural semantics\/}, functions imported into and exported from
typed modules are wrapped in proxies that enforce the static type discipline
with run-time checks and track responsibilities~\citep{tf-popl-2008,
tfffgksst-snapl-2017}. Thus, when {\tt pack} is imported into a typed module,
the run-time system checks that it is a function and wraps it in a protective
proxy, which in turn, enforces the type of the function result with run-time checks.
Analogously the run-time system wraps each exported function of a typed module
such as {\tt crypto-pack} in a proxy that checks its arguments.  These checks
protect functions exported from typed modules against applications to
wrong arguments in untyped code.

As this analysis implies, if a return-type check fails, the problem is with
the untyped module that supplied an incorrect function, here {\tt pack-lib}. If
an argument-type check fails, responsibility lies with {\tt client}. Due to the
proxies wrapped around functions, {\it Natural\/} can easily track the responsible party
for each check. Thus, in the example of figure~\ref{fig:example}, as
\texttt{pack} returns, the return-type check fails and Natural blames
the boundary between \texttt{pack-lib} with {\tt typed-pack-lib} and
{\tt crypti-packed-lib}, respectively, for the two {\tt define}s in {\tt client}.

@@

In order to locate the responsible party for failed checks, Transient
maintains a map from values to the boundaries between typed and untyped modules that
they cross, plus the corresponding types. In the example, the map records that
\texttt{pack} crosses from \texttt{pack-lib} to {\tt typed-pack-lib} and
from {\tt pack-lib} to {\tt crypto-pack-lib} with the type that appears in the
{\tt required/typed} forms in the example. Since the failed check corresponds to
the return type of \texttt{pack}, Transient blames the sources of the two
boundary crossings, which in this particular case is \texttt{pack-lib}. In
general though, Transient blames more than one party. In fact, the theoretical
work of~\citet{gfd-oopsla-2019} shows that for some programs Transient
constructs a blame sequence that excludes responsible parties and includes modules
irrelevant to the failing check.

Under {\it Erasure\/}, the compiler checks the specified types and
then discards them when it generates code. The generated code includes run-time
checks that ensure the dynamic safety of all operations as specified in the
underlying untyped language. Hence, neither the call to {\tt typed-pack} nor the
call to {\tt crypto-pack} signals an error due to the gradual type system. If
at some later point {\tt client} tries to inspect the elements of the lists
that \texttt{typed-pack} and {\tt crypto-pack} are supposed to produce, Racket's
safety checks signal a violation and point to some place in {\tt client}. The
information in this exception, plus its stack trace, may help the programmer
find the source of the impedance mismatch between the specified return types of
{\tt pack} in the two typed modules and the actual return value of {\tt pack}.

The following table summarizes the illustration. Each cell describes the result
of evaluating the column's definition (in {\tt client}) under the row's
semantics.
%% -----------------------------------------------------------------------------
\begin{center}
  \begin{tabular}{l|l@{\qquad\quad}l}
                  & {\tt public-data}              & {\tt secret-data}                                 \\
\hline %% ----------------------------------------------------------------------------------------
{\it Natural\/}   & error, blaming the boundary between & error, blaming the boundary between          \\
                  & {\tt pack-lib} and {\tt typed-pack-lib} & {\tt pack-lib} and {\tt crypto-pack-lib} \\
                  &                                &                                                   \\
{\it Transient\/} & no error                       & error, blaming the boundary crossings             \\
                  &                                & \quad {\tt pack-lib} to {\tt typed-pack-lib}, and \\
                  &                                & \quad {\tt pack-lib} to {\tt crypto-pack-lib}     \\
                  &                                &                                                   \\
{\it Erasure\/}   & no error*                      & no error*                                         \\
                  &   \multicolumn{2}{|c}{*but {\em Erasure} does signal an error on list access}
\end{tabular}
\end{center}
%% -----------------------------------------------------------------------------

>>>>

\end{description}

The mistaken comment in {\tt pack-lib} causes an impedance mismatch, with which each
of the three semantics deals differently.  Under {\it the Natural semantics\/},
functions imported into and exported from typed modules are wrapped in proxies that
enforce the static type discipline with run-time checks and track
responsibilities~\citep{tf-popl-2008, tfffgksst-snapl-2017}. Thus, when {\tt pack}
is imported into a typed module, the run-time system checks that it is a function
and wraps it in a protective proxy, which in turn, enforces the type of the function
result with run-time checks.  Analogously the run-time system wraps each exported
function of a typed module such as {\tt crypto-pack} in a proxy that checks its
arguments.  These checks protect functions exported from typed modules against
applications to wrong arguments in untyped code.

As this analysis implies, if a return-type check fails, the problem is
that the untyped module supplied a function, here {\tt
pack-lib}, that is not a  match for the type ascribed by the typed module. 
Hence either the type at the boundary between the two modules is wrong or, 
if the programmer trusts the type, the untyped module is at fault.
If an argument-type check fails, responsibility  lies with
{\tt client}. After all either the type it ascribes to the argument is
wrong or the argument it produced clashes with the type. Due to 
proxies, {\it Natural\/} can easily track the 
boundary, type and responsible parties that correspond to each check. 
Thus, in the example of figure~\ref{fig:example}, as
\texttt{pack} returns, the return-type check fails and Natural blames
the boundary between \texttt{pack-lib} with {\tt typed-pack-lib} and
{\tt crypto-pack-lib}, respectively, for the two {\tt define}s in {\tt client}.

@@

In order to locate the corresponding boundaries for failed checks, Transient
maintains a map from values to the boundaries between typed and untyped modules that
they cross, plus the corresponding types. In the example, the map records that
\texttt{pack} crosses from \texttt{pack-lib} to {\tt typed-pack-lib} and
from {\tt pack-lib} to {\tt crypto-pack-lib} with the type that appears in the
{\tt required/typed} forms in the example. Since the failed check corresponds to
the return type of \texttt{pack}, assuming that the type is correct, 
the responsible parties are the sources of the two
boundary crossings, \texttt{pack-lib}. In
general though, Transient blames more than one boundary. In fact, the theoretical
work of~\citet{gfd-oopsla-2019} shows that for some programs Transient
constructs a blame sequence that excludes responsible parties and includes modules
irrelevant to the failing check.

Under {\it Erasure\/}, the compiler checks the specified types and
then discards them when it generates code. The generated code includes run-time
checks that ensure the dynamic safety of all operations as specified in the
underlying untyped language. Hence, neither the call to {\tt typed-pack} nor the
call to {\tt crypto-pack} signals an error due to the gradual type system. If
at some later point {\tt client} tries to inspect the elements of the lists
that \texttt{typed-pack} and {\tt crypto-pack} are supposed to produce, Racket's
safety checks signal a violation and point to some place in {\tt client}. The
information in this exception, plus its stack trace, may help the programmer
find the source of the impedance mismatch between the specified types of
{\tt pack} in the two typed modules and its actual results.

The following table summarizes the illustration. Each cell describes the result
of evaluating the column's definition (in {\tt client}) under the row's
semantics.
%% -----------------------------------------------------------------------------
\begin{center}
  \begin{tabular}{l|l@{\qquad}l}
                  & {\tt public-data}              & {\tt secret-data}                                 \\
\hline %% ----------------------------------------------------------------------------------------
{\it Natural\/}   & error, blaming the boundary between & error, blaming the boundary between          \\
                  & {\tt pack-lib} and {\tt typed-pack-lib} & {\tt pack-lib} and {\tt crypto-pack-lib} \\
\hline
{\it Transient\/} & no error                       & error, blaming the boundaries between            \\
                  &                                & {\tt pack-lib}  and {\tt typed-pack-lib}, and \\
                  &                                & {\tt  pack-lib} and {\tt crypto-pack-lib}     \\
\hline
 {\it Erasure\/}   & no error*                      & no error*                                         \\
                  &   \multicolumn{2}{|c}{*but {\em Erasure} does signal an error on list access}
\end{tabular}
\end{center}
%% -----------------------------------------------------------------------------

### 2021-05-25 writing/rational-natural.tex
clarifying blame and boundaries + a small edit about narrow vs broad blame information

\mf{Technically a Natural blame trail could be a Natural exception trail. Hmph? }
\mf{I added a sentence below this definition and the next to address this issue.}

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\blame{\system, \conf_i}\}$      & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$  & otherwise \\ %
\end{tabular} \right. %
\]
where $\blame{\system, \conf}$ denotes the module (of $\system$) that $\conf$ blames,
and $\exception{\system, \conf}$ denotes the first untyped location in the stacktrace produced by $\conf$.
\end{quote}
Note how this definition interprets ``blame information'' broadly, as
any information from a failed run-time check in a blame-oriented semantics.

>>>>

\mf{Technically a Natural blame trail could be a Natural exception trail. Hmph? }
\mf{I added a sentence below this definition and the next to address this issue.}
\cd{I revised a bit.}

Here is a rigorous definition. 
\begin{quote}
\it A \emph{Natural blame trail} is a sequence of scenarios $\conf_0,...\conf_n$ of
a program $\system$ such that for all $0 \leq i \leq n - 1$, $\conf_i \subset
\conf_{i+1}$ and
\[ \conf_{i+1} \setminus \conf_i = %
\left\{ %
\begin{tabular}{ll} %
$\{\blame{\system, \conf_i}\}$      & if $\conf_i$ produces blame \\ %
$\{\exception{\system, \conf_i}\}$  & otherwise \\ %
\end{tabular} \right. %
\]
where $\blame{\system, \conf}$ denotes the module (of $\system$) that $\conf$ blames,
and $\exception{\system, \conf}$ denotes the first untyped location in the stacktrace produced by $\conf$.
\end{quote}
Note how, in the absence of ``blame information'' in the narrow sense, 
this definition interprets ``blame information'' broadly, as
any information from a failing run-time check.

### 2021-05-25 writing/rational.tex
Small fixes from meeting last week
clarifying blame and boundaries + a small edit about narrow vs broad blame information

Section~\ref{sec:why-rational} explains how a migratory type setting helps with
finding the source of an impedance mismatch. Roughly speaking, it encourages the
rational programmer to equip a module with types if it is blamed in an error
message. A sequence of such steps makes up a path in the
lattice of type migration~\citep{tfgnvf-popl-2016}. The lattice serves as the
universe \mf{what does this sentence really mean? what does it contribute?} for the definition modes of the rational programmer (see
sec.~\ref{sub:stuff}).

>>>>

Section~\ref{sec:why-rational} explains how a migratory type setting helps with
finding the source of an impedance mismatch. Roughly speaking, it encourages the
rational programmer to equip a module with types if it is blamed in an error
message. A sequence of such steps makes up a path in the
lattice of type migration~\citep{tfgnvf-popl-2016}. The lattice describes the
space in which the modes of the rational programmer search for bugs (see
sec.~\ref{sub:stuff}).
%\mf{what does this sentence really mean? what does it contribute?}
%\ll{Addressed}
